---
title: "Working with phenotypic data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This notebook is delivered "As-Is". Notwithstanding anything to the contrary, DNAnexus will have no warranty, support, liability or other obligations with respect to Materials provided hereunder.

[MIT License](https://github.com/dnanexus/UKB_RAP/blob/main/LICENSE) applies to this notebook.

# Initializing `renv`

Start by installing `renv` library and initializing `renv` project. The `renv` initialization creates a new `renv.lock` file where the R version and the list of used packages with their version are tracked and also a new `renv` folder containing settings file and script to activate virtual environment.

```{r}
#install.packages('renv')
renv::init()
```

## Installing libraries

Uncomment the install commands if you are comfortable with the library license and want to install and run the parts notebook that depend on the library.

```{r}
#renv::install('tidyverse')
#renv::install('Hmisc')
#renv::install('rlang')
```

# Transfer files from UKB RAP Project

Following code assumes that you have `data.csv` file in the root directory of your current UKB RAP project. You can create this file by using Table exporter.

```{r}
system('dx download data.csv')
```

# Make analysis

```{r}
library('tidyverse')
library('Hmisc')
```

Load phenotypic data.

```{r}
data <- read.csv('data.csv')
head(data)
```

Plot histogram for each column.

```{r}
ggplot(gather(data), aes(value)) +
  geom_bar(stat = 'count', binwidth = 0.5) +
  facet_wrap(~key, scales = 'free_x') + theme(axis.text.x=element_text(angle = 45, vjust = 0.5))
```

Print descriptive statistics.

```{r}
describe(data)
```

Extract rows that have `Age at the recruitment` age less than 5th percentile.

```{r}
data_young <- data %>% filter(p21022 < quantile(p21022, 0.05, na.rm = TRUE))
```

Look descriptive statistics of subset data.

```{r}
describe(data_young)
```

Write subset data frame as CSV file.

```{r}
write_csv(data_young, 'age_month_birth_young.csv')
```

# Snapshot installed libraries

Save the state of the project library to the lockfile (called `renv.lock`).

```{r}
renv::snapshot()
```

# Upload RStudio project into UKB RAP Project storge

```{r}
system('dx-backup-folder -d /.Backups/rstudio_webinar.tar.gz')
```
