---
title: "Installing R and Bioconductor Packages using `renv`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This notebook is delivered "As-Is". Notwithstanding anything to the contrary, DNAnexus will have no warranty, support, liability or other obligations with respect to Materials provided hereunder.

[MIT License](https://github.com/dnanexus/UKB_RAP/blob/main/LICENSE) applies to this notebook.

# Using `renv`

In this markdown project, we'll install some packages reproducibly with `renv`, which will save us the headache of reinstalling packages from scratch everytime.

There are the following steps for using `renv` on UKB RAP:

1) Create our project
2) Install `renv`, use `renv::init()`
3) Install our packages of interest, then call `renv::snapshot()`
4) Zip our project up, and upload the zip file to our UKB RAP project using `dx upload`

To restore our project, we need to do the following:

a) upload our project zip file from project storage using `dx download`
b) unzip our project
c) open the project by clicking on the `my_project.rproj` file.

## 1. Create Our Project

Open this RMarkdown file outside of a project (If you're in one, click "Files >> Close Project").

The first thing we'll do is create a project called `my_project`. We'll do this with `usethis::create_project()` (after installing it, of course).

```{r}
install.packages("usethis")
usethis::create_project("my_project")
```

## 2. Install `renv`

`renv` will allow us to make our project reproducible by creating a private project library.

We start up our project environment by using `renv::init()`. This will set up the file structure for the private library.
Uncomment the install commands if you are comfortable with the library license and want to install and run the parts notebook that depend on the library.

```{r}
#install.packages("renv")
renv::init()
```

# 3. Install our packages

Now we can install our packages into our project, including Bioconductor packages.

When we're done, we call `renv::snapshot()`, which scans the libraries we've installed and updates the `renv.lock` file.

```{r}
#install.packages("skimr")
#BiocManager::install(c("GenomicFeatures", "AnnotationDbi"))
renv::snapshot()
```

## Do some work



```{r}

```

## 4. Zipping our project

When we're done, we can run the following in our project to zip up our project and use `dx upload` to upload the RStudio Project to our UKB RAP Project:

```{r}
system("dx-backup-folder -d /.Backups/my_project.tar.gz")
```

# Restore Project and Dependencies

## A. Download and Unzip

Now you can download your RStudio project from project storage, unzip it, and open the project up again by clicking on the `my_project.rproj` file in the project.

We do this using `dx-restore-folder`, which will let us restore the folder we've stored in `/.Backups/my_project.tar.gz`).

```{r}
system("dx-restore-folder /.Backups/my_project.tar.gz")
```

## B. Open our project up

In the RStudio Files pane, you can then open the project by opening the `.Rproj` file and have access to all of your libraries.
