---
title: "Rmarkdown On UKB RAP"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document will show you how to access files from RAP 

## List files in current project

We'll need to figure out where our files are in the RAP project storage.

```{r}
system("dx ls 'Showcase metadata'")
```
## Load the file using dxFUSE

The first way we can access a file in a project is by prepending a `/mnt/project/` to our file path. This lets us access files in our project. 

```{r}
field_data2 <- read.delim("/mnt/project/Showcase metadata/field.tsv")
```

```{r}
head(field_data2)
```

## Alternative: download file into Workspace Storage

If we want to be explicit in downloading files, we can utilize `dx download` within `system()` to download our files from the DX Toolkit.

Note that we 

```{r}
system("dx download '/Showcase metadata/field.tsv'")
list.files()
```

## Load the file into R

```{r}
field_data <- read.delim("field.tsv")
```

```{r}
head(field_data)
```


## Writing our file to the RStudio Workspace

```{r}
write.csv(field_data2, file="updated_field_data.csv")
list.files()
```

## Saving Our file back into Project Storage

```{r}
system("dx upload updated_field_data.csv")
```

## Save our notebook into Project storage as well

```{r}
system("dx upload ukb_test.Rmd")
```